---
title: "Assignment-01"
author: "mike-lid", "JebP"
date: "9/6/2021"
output: html_document
---

```{r load packages, message=FALSE}
#above line labels the code chunk as "load packages" without seeing any of the package loading messages
  library (sf)
  library(tidyverse)
  library(ggspatial)
  library(ggthemes)
```

```{r load packages, message=FALSE}
#the following data includes commercial parking locations in Cambridge, polygon data of metered parking spots, and boundaries for open space. We are curious to see the relationship between available Cambridge parking and locations of open/green space. Comparing the geographic distribution of space for parked cars versus space for people, plants, and animals may yield interesting insights into land use priorities. We hope that by writing "message=false", that this chunk will not appear in the knit

  commercialparking <- st_read("CommParking.geojson")
  meters <- st_read("MetParking.shp.zip")
  openspace <- st_read ("OpenSpace.shp.zip")
```

#This first map uses a high contrast dark theme. Open space polygons and commercial parking points pop from the page, while the metered parking more subtly suggests denser corridors, where people will pay to park on the street. 
```{r}
ggplot() +
  
ggtitle("Cambridge Park(ing)") +
  
#BASE MAP
    annotation_map_tile(
    zoomin = 0, 
    progress = "none", 
    type = "cartodark") +
    geom_sf() +
    labs(caption = "Map tiles and data by OpenStreetMap") +
  
#DATA 
   geom_sf(
    data = openspace, 
    color = NA, 
    alpha = 0.6,
    aes(fill = "Open Space")
    ) +
  geom_sf(
    data = meters, 
    fill = "white",
    aes(color = "Parking Meters")
    ) +
  geom_sf(
    data = commercialparking, 
    shape = 20, 
    size = 3, 
    aes(color = "Commercial Parking")
    ) +
  scale_color_manual(values = c("deeppink", "darkviolet"), name = "") +
  scale_fill_manual(values = "chartreuse", name = "") +
  
  theme_void()

  
```

#This next map uses a toned-down palette. The simple basemap allows the viewer to focus on the point and polygon layers atop it. Changing the symbology of the metered parking polygons to have just a fill with no outline shows the true size of the parking spaces, but makes them a bit hard to find. 
```{r}
ggplot() +
  
ggtitle("Cambridge Park(ing)") +
  
#BASE MAP
    annotation_map_tile(
    zoomin = 0, 
    progress = "none", 
    type = "cartolight") +
    geom_sf() +
    labs(caption = "Map tiles by OpenStreetMap; data from City of Cambridge") +
  
#DATA 
   geom_sf(
    data = openspace, 
    color = NA, 
    alpha = 0.5,
    aes(fill = "Open Space")
    ) +
  geom_sf(
    data = meters, 
    color = NA,
    shape = NA,
    aes(fill = "Metered Parking")
    ) +
  geom_sf(
    data = commercialparking, 
    shape = 3, 
    size = 2, 
    fill = NA,
    aes(color = "Commercial Parking")
    ) +
  
  scale_fill_manual(values = c("brown3", "forestgreen"), name = "") +
  scale_color_manual(values = "brown1", name = "") +
  
  theme_void()
```


#To test a new representation of the metered parking spaces, we derived the centroids of each metered parking polygon. This means that metered parking now appears as points, like the commercial parking layer. The orange color distinguishes metered parking from commercial parking, while keeping them in the same visual category (i.e. warm color = parking). For an unknown reason, the centroid points become black and deformed when a basemap was added, so I removed the basemap and reinserted coordinate marks for context. 

```{r}

#finding centroids of metered parking polygons
ggplot(st_centroid(meters)) +
  geom_sf(
    size = 0.1,
    shape = 20,
    aes(color = "Metered Parking"),
    ) +

#TITLE
  ggtitle("Cambridge Park(ing)") +  
  
  
#open space and commercial parking 
   geom_sf(
    data = openspace, 
    color = NA, 
    aes(fill = "Open Space"),
    ) +
  geom_sf(
    data = commercialparking, 
    shape = 20, 
    size = 3, 
    aes(color = "Commercial Parking"),
    ) +
  
#LEGEND  
  scale_color_manual(values = c("brown1", "darkorange"), name = "") +
  scale_fill_manual(values = "forestgreen", name = "") +
  
   theme(panel.background = element_rect(fill = "white"),
        legend.key = element_rect(fill = "white"))

```

#Finally, this ugly version overdoses on color. The basemap distracts from the other layers and it is hard to distinguish what is important. The blue open space polygons might be confused with water. The commercial parking points are too large and the star shape is needlessly complicated. The metered parking polygons are almost invisible. 
```{r}
ggplot() +
  
ggtitle("Cambridge Park(ing)") +
  
#BASE MAP
  
    annotation_map_tile(
    zoomin = 0, 
    progress = "none", 
    type = "stamenwatercolor") +
    geom_sf() +
    labs(caption = "Map tiles by Stamen Design. Data by OpenStreetMap") +

 
#DATA 
   geom_sf(
    data = openspace, 
    color = NA, 
    alpha = 0.5,
    aes(fill = "Open Space")
    ) +
  geom_sf(
    data = meters, 
    color = NA,
    shape = NA,
    aes(fill = "Metered Parking")
    ) +
  geom_sf(
    data = commercialparking, 
    shape = 11, 
    size = 5, 
    fill = NA,
    aes(color = "Commercial Parking")
    ) +
  
  scale_fill_manual(values = c("brown1", "blue"), name = "") +
  scale_color_manual(values = "brown1", name = "") +
  
  theme_void()
```



